### 2021-05-19 project_ashish_monograph/PCC.tex
monograph updated

Specific techniques are employed for the reconstruction of clusters in the pixel detector and all the reconstructed clusters are used for counting to calculate the instantaneous luminosity. Algorithm used for finding clusters check for pixels whose signal to noise ratio is more than 6 and then merge adjacent pixel. These algorithms estimates cluster position in two directions and cluster charge. A cluster is a set of adjacent pixels and only pixels above certain minimum value of charge are considered. Position of a cluster in X and Y directions are obtained by fitting X and Y projections to templates that are estimates for cluster shapes \cite{Chatrchyan:2014fea}.

>>>>

LHC 2018 run can be categorized as follows, \\

1. Run2018A \\

Run number ranges from 315255 to 316995 \\
Number of runs per period is 146 \\

2. Run2018B\\

Run number ranges from 317080 to 319311\\
Number of runs per period is 132\\

3. Run2018C \\

Run number ranges from 319337 to 320065\\
Number of runs per period is 88 \\

Modules which are observed to function efficiently are considered for luminosity calculation. The module veto lists correspond to second Run2018D and select each detector part L2, L3, L4, D1, D2 and D3. Barrel layer 1 (L1) is not used for luminosity determination \cite{vetolist}. \\

1. veto lateRunD lowcut tight   (Number of vetoed modules 1387) \\
2. veto lateRunD lowcut tight B1 (Number of vetoed modules 1850)\\
3. veto lateRunD lowcut tight B2 (Number of vetoed modules 1766)\\
4. veto lateRunD lowcut tight B3 (Number of vetoed modules 1641)\\
5. veto lateRunD lowcut tight F1 (Number of vetoed modules 1813)\\
6. veto late RunD lowcut tight F2 (Number of vetoed modules 1799)\\
7. veto late RunD lowcut tight F3 (Number of vetoed modules 1800)\\

Specific techniques are employed for the reconstruction of clusters in the pixel detector and all the reconstructed clusters are used for counting to calculate the instantaneous luminosity. Algorithm used for finding clusters check for pixels whose signal to noise ratio is more than 6 and then merge adjacent pixel. These algorithms estimates cluster position in two directions and cluster charge. A cluster is a set of adjacent pixels and only pixels above certain minimum value of charge are considered. Position of a cluster in X and Y directions are obtained by fitting X and Y projections to templates that are estimates for cluster shapes \cite{Chatrchyan:2014fea}. \\

### 2021-05-19 project_ashish_monograph/bkg.tex
references/figures updated
monograph updated

\newpage \section{Backgrounds}
\label{sec:bkg}
(background in vdM, afterglow backgrounds in Physics, subtraction of these) \\

N= \sigma \:L + N_{background} \\

L = $ \int L_{inst} \: dt = \int \frac{R(t)}{\sigma_{vis}} \:dt $  \\

@@

N: total number of event including signal and background. \\

@@

@@

s: center of mass energy of colliding protons. \\

@@

The pixel cluster counting (PCC) method produces a small out-of-time response (OOTR, or afterglow) that has two types.\\

Type 1: fake hits in the bunch crossing after the colliding bunches (signal charge spillover).\\

Type 2: Activated material in the detector due to the large radiation doses, exponentially decays for many bunch crossings.\\

Afterglow effects are tiny in the case of tracklets (2x and 3x coincidences).\\

Due to both the late-arriving particles and energy originating from activated detector material, the pixel data need to be corrected for afterglow effects. This effect was studied in data taken using random triggers (triggers that fire randomly in bunch crossings spread over the whole LHC orbit except for the abort gap), considering the pixel cluster counts in bunch crossings after bunch trains. The resulting correction depends on the LHC filling scheme and for a typical 2011 fill with 1380 bunches corresponds to a subtraction of close to 2.8$\%$ of the integrated luminosity per luminosity section. \\

The afterglow corrections for PCC during 2017 run are in the range 2-5 $\%$ for the Type 1 corrections and 2-3 $\%$ averaged over all active bunches for the Type 2 corrections.

@@

Fig. 7 shows the afterglow noise induced by one colliding bunch. Since it is normalized to 1 at the first bin it means the first bin is the colliding bunch. The values at bins $>$ 1 are the cluster counts observed for bunches after the colliding bunch divided by the cluster counts observed by the colliding bunch. There are three parts to this distribution: \\

bin 1: colliding bunch \\

bin 2 (empty bunch): bunch after colliding bunch, noise observed is due to electronics signal on same pixels from colliding bunch. This is called Type 1. \\

bins $>$ 2 (empty bunches): noise observed is due to "albedo" which is material activated by the radiation of the collisions, it decays exponentially with time just like radioactive material. It produces secondary particles which create hits late in time. This is called Type 2.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\columnwidth]{./afterglow.png}
  \caption{Results of the pixel afterglow 'deconvolution fit'}
  \label{fig:LHC}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\columnwidth]{./PCC_background.png}
  \caption{ \onehalfspacing \cite{}.}
  \label{fig:CMS}
\end{figure}

>>>>

\section{Backgrounds}
\label{sec:bkg}
(background in vdM, afterglow backgrounds in Physics, subtraction of these) \\

The cluster rate equation is modified in the presence of background events. \\

N = \sigma \:L_{int} + N_{background} \\

L_{int} = $ \int L_{inst} \: dt = \int \frac{R(t)}{\sigma_{vis}} \:dt $  \\

@@

N: total number of signal and background events. \\

@@

$L_{int}$: integrated luminosity. \\

@@

s: center-of-mass energy of colliding protons. \\

@@

$N_{background}$ need to be subtracted during vdM scans for precise determination of $\sigma_{vis}$. During high pileup physics runs, background contribution to instantaneous luminosity/$\sigma_{vis}$ based on the pixel cluster counting (PCC) method arises from a small out-of-time response (OOTR, or afterglow) that has two types.\\

Type I: fake hits in the bunch crossing after the colliding bunches (signal charge spillover).\\

Type II: Activated material in the detector due to the large radiation doses, exponentially decays for many bunch crossings.\\

Fig. 11 shows the afterglow noise induced by one colliding bunch. Since the distribution is normalized to 1 at the first bin it means the first bin is the colliding bunch. The values at bins $>$ 1 are the cluster counts observed for bunches after the colliding bunch divided by the cluster counts observed by the colliding bunch. There are three parts to this distribution: \\

(i) bin 1 is the colliding bunch \\

(ii) bin 2 (empty bunch) is the bunch after colliding bunch, noise observed is due to electronics signal on same pixels from colliding bunch. This is Type I afterglow effect. \\

(iii) bins greater than 2 are all empty bunches. Noise observed is due to "albedo" which is material activated by the radiation of the collisions, it decays exponentially with time just like radioactive material. It produces secondary particles which create hits late in time. This is called Type II afterglow effect. \\

 Afterglow effects for various bunches in a beam is shown in Fig. 12 where the intial rising curve correspong to Type I and decay curve correspond to Type II afterglow effect. It was studied in data taken using random triggers (triggers that fire randomly in bunch crossings spread over the whole LHC orbit except for the abort gap), considering the pixel cluster counts in bunch crossings after bunch trains. The resulting correction depends on the LHC filling scheme and for a typical 2011 fill with 1380 bunches corresponds to a subtraction of close to 2.8$\%$ of the integrated luminosity per lumi section. The afterglow corrections for PCC during 2017 run are in the range 2-5 $\%$ for the Type I corrections and 2-3 $\%$ averaged over all active bunches for the Type II corrections \cite{Sirunyan:2759951}.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\columnwidth]{./PCC_background.png}
  \caption{ \onehalfspacing \cite{}.}
  \label{fig:CMS}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\columnwidth]{./afterglow.png}
  \caption{Results of the pixel afterglow fit.}
  \label{fig:LHC}
\end{figure}

### 2021-05-19 project_ashish_monograph/syst.tex
monograph updated

\newpage
\textbf{Systematics uncertaintities during physics runs (high pileup)}:\\

6. Non-Linearity: extrapolation of $\sigma_{vis}$ from vdM scans to high pileup physics data-taking periods can cause it to vary with per bunch instantaneous luminosity and pileup if the relation between luminosity and pileup is not linear.                               \\

>>>>

\textbf{Systematics uncertaintities during physics runs (high pileup)}:\\

6. Non-Linearity: extrapolation of $\sigma_{vis}$ from vdM scans to high pileup physics data-taking periods can cause $\sigma_{vis}$ to vary with per bunch instantaneous luminosity and pileup if the relation between luminosity and pileup is not linear.                               \\

### 2021-05-19 project_ashish_monograph/vdM.tex
monograph updated

\newpage \section{van der Meer calibration}
\label{sec:vdm}
Van der Meer (vdM) scans are done to calculate the visible cross section $\sigma_{vis}$ of the pixel detector which is a calibration constant in the relation between rate of a quantity (clusters, coincidences) read out by the detector and the instantaneous luminosity. \\

@@

\newpage Thus, the expression for instantaneous luminosity is given by \\

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\columnwidth]{./vdm_image.png}
  \caption{\onehalfspacing Figure showing two LHC bunches in green and blue approaching each other and colliding giving rise to an overlapping region showing in red whose area is determined during vdM scans.}
  \label{fig:CMS}
\end{figure}

>>>>

 \section{van der Meer calibration}
\label{sec:vdm}
Van der Meer (vdM) scans are done to calculate the visible cross section $\sigma_{vis}$ of the pixel detector which is a calibration constant in the relation between rate of a quantity (clusters, coincidences) read out by the detector and the instantaneous luminosity. \\

@@

Thus, the expression for instantaneous luminosity is given by \\

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\columnwidth]{./vdm_image.png}
  \caption{\onehalfspacing Figure showing two LHC bunches in green and blue approaching each other and colliding giving rise to an overlapping region shown in red whose area is determined during vdM scans.}
  \label{fig:CMS}
\end{figure}
